%------------------------------------------------------------------------------
% Syneight - A soft-realtime transaction monitor.
% Copyright (C) 2003-2004 The Syneight Group.
%
% TODO: License.
%------------------------------------------------------------------------------

\chapter{Motivation}


%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
\section{What is \SYNEIGHT}
Leading idea: ``Abstract description of \SYNEIGHT''.

Pure prose.

In short:
\begin{description}
\item[Domain independent] define -- use C++ Network Programming,
Schmidt, page 8

``Common middleware services augment distribution middleware by defining
high-level domain independent services, such as event notification,
logging, persistence, security, and recoverable transactions. Whereas
distribution middleware focuses largely on managing end-system resources
in support of an object-oriented distributed programming model, common
middleware services focus on allocating, scheduling, and coordinating
various resources throughout a distributed system. Without common
middleware services, these end-to-end capabilities would have to be
implemented ad hoc by each networked application.''

\item[Transaction oriented] use the term extended atomic transactions
\item[Soft real-time] define -- Kopetz
\item[Cluster / SMP] SMP: uniform memory, cluster. Clustering for
scalability and availability. We use SMP machines as a nodes since to
keep the number of nodes small -- and thus, the communication
overhead.
\end{description}

Think about ordering of these issues. 

Problem: Not \MMORGS, not \NVES. These two are motivations which we
will discuss in great length in this chapter -- but they are not
defining \SYNEIGHT. Thus, they do not appear in this section.

10 Sentences

%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
% OLD VERSION OF THIS SECTION
%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
%%\section{What is \SYNEIGHT}
%%
%%\SYNEIGHT is a transaction and object oriented middleware which
%%empathizes SMP-clustering and soft-realtime processing. Originally,
%%\SYNEIGHT was motivated by networked virtual environments
%%(\NVES), in particular by large-scale simulations developed by the
%%entertainment industry.
%%%
%%Currently, so-called Massively Multiplayer Role Games (\MMORGS) are
%%gaining more and more importance in this industry (ADDREF). The notion
%%of \MMORGS will be explained in detail in \vref{sec:I:motivation:mmorgs}.
%%%
%%State of the art \MMORGS are challenging in various ways. While the
%%gaming industry used to adopt (at least partially) an ad-hoc style of
%%development for traditional game projects, these applications require a
%%much more rigorous approach.
%%%
%%
%%%%% A \MMORG is demanding wrt.~scalability, reliability, modifiability. 
%%%%% 
%%%%% 
%%%%% These simulations are challenging in various ways. Scalability,
%%%%% modifiability, extensibility, 
%%%%% are among the most the
%%%%% important goals.
%%
%%
%%%
%%% scalability
%%% modifiability
%%%    extensibility
%%
%%% dependability reliability fault tolerance
%%
%%% realtime
%%
%%
%%% leeway for experimentation
%%% inherent game logic (nothing should be unmotivated)
%%%    seamlessness 
%%
%%%
%%% development of A mmorg -- quest for a domain-independent base, to
%%% ensure modifiability, extensibility... together with the need for a
%%% uniform interface formulation gave raise to the idea of
%%% \SYNEIGHT. 
%%
%%In this chapter, we will review \NVES, \MMORGS and the currently
%%available technology. Based upon this outline, we will argue, that a
%%new approach is necessary to solve the inherent problems of this
%%application domain. 
%%%
%%This domain comes with a broad and general set of
%%requirements, which cannot be solved independently from each other.
%%% --> against tool assembly
%%Furthermore, we regard a generalized approach as necessary to ensure
%%the required flexibility and modifiability. Fortunately, \NVES allow
%%to obtain a condensed and abstract set of fundamental
%%requirements. These requirements form the starting point of the
%%\SYNEIGHT middleware which is projected to serve as such an abstract
%%basis for \NVES in terms of a domain-independent middleware.
%%
%%%ranging from classical software engineering goals such
%%%as modifiability and extensibility to more specific requirements such
%%%as soft realtime processing and 
%%
%%% no throwing together of existing tools
%%% stability (wrt. interface, internals, ...) by orthogonalization
%%% (i.e., generalization)
%%% 
%%
%%%% 
%%%% 
%%%% 
%%%% \SYNEIGHT is a middleware, so before we can tell what \SYNEIGHT exactly is
%%%% we have to explain what a middleware is. Wikipedia defines middleware in
%%%% the following way:
%%%% 
%%%% \begin{definition} 
%%%% Middleware in computing terms is used to describe a software agent
%%%% acting as an intermediary, or as a member of a group of intermediaries,
%%%% between different components in a transactional process.  The classic
%%%% example of this is the separation which is attained between the client
%%%% user and the database in a client/server situation.  The reason for
%%%% introducing middleware in such a situation is to better service client
%%%% requests by reducing the number of resource-expensive connections to the
%%%% database and more efficiently passing the requested data back.  
%%%% \end{definition}
%%%% 
%%%% So in short it is an API or software collection that hides complexity.
%%%% 
%%%% There are two kinds of middleware. Domain dependant and domain in
%%%% dependant middleware. Domain dependant means that it is designed for a
%%%% specific application and can be used only for one application or in best
%%%% case for only one specific kind of applications. In contrast domain
%%%% independent middleware can be used by many different kinds of
%%%% applications. \SYNEIGHT is the second. A domain independent middleware. 
%%%% 
%%%% But what 
%%%% 
%%%More specifically \SYNEIGHT is a domain independent middleware. That
%%%means that it can be used in a variety of domains like virtual
%%%environments, entertainment, education, cooperative work or science. In
%%%contrary to that there are also domain dependant middleware. This kind
%%%of middleware would be usable only for one specific kind of application
%%%like military simulations.
%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
% END OF THE OLD VERSION OF THIS SECTION
%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%

%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
\section{Motivating Application Domain}
Leading idea: ``Which domain, why is the domain interesting, which
domain structure.''

Better Structure (convert this section here into this form):
\begin{itemize}
	\item MMORGS -- A new Entertainment World
	\begin{itemize}
		\item Define MMORGS
		\item ``Why MMORGS''
		\item short: the rest of the section
	\end{itemize}
	\item ''what are nves'' with applications
	\item MMORGS in detail
	\begin{itemize}
		\item maybe more detailed definition
		\item MMORGS are a special case of NVEs
		\item industrial interest
  		\item characteristics \& challenges
		\item feasibility study
	\end{itemize}
\end{itemize}

%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%% 
\subsection{\MMORGS -- A new Entertainment World}
Needs a better catchy expression:
\begin{itemize}
	\item A new Entertainment World
	\item The new Challenge
	\item \dots
\end{itemize}

\subsubsection{What are \MMORGS}
Definition \ldots

\subsubsection{The interest in \MMORGS}
\begin{itemize}
	\item Academic Interest
	\begin{itemize}
		\item Subsumes Many general NVE Problems
		\item It is a game:
		\item Incredible wishlist
		\item Tolerant ( The world is not going under when something crashes )
		\item Bottomline: Interesting research ground
	\end{itemize}
	\item Industrial Interest
	\begin{itemize}
		\item Feasibility study
	\end{itemize}
\end{itemize}

%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
\subsection{What are \NVES}

\MMORG is a specialization of \NVE.

based on virtual environments papers

adding networking (cf. overview paper) -- usage of networking in terms
of a, load distribution, and b, geographically dispersed access

read the stuff! /usr/local/data/syneight/literature/nve/survey, read
the papers, check out what's going on.

Also integrate the applications:
\begin{itemize}
        \item Military Simulations
        \item CSCW ( Computer Supported Cooperative Work for example
              Virtual meetings)
        \item Educational
        \item Entertainment
        \item \dots
\end{itemize}

Application Elements? Leave blank in the first run.
\begin{itemize}
        \item X-Reality
        \item Augmented Reality
        \item Simulation
        \item \dots
\end{itemize}

%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
\section{\MMORGS in detail}
Characteristics \& Challenges
\begin{itemize}
	\item maybe more detailed definition
	\item MMORGS are a special case of NVEs
	\item industrial interest
	\item characteristics \& challenges
	\item feasibility study
\end{itemize}

%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
\section{The services of a \MMORG}
Leading Idea: ``OK. So that's the way you can structure such a
biest.''

\textbf{Feasibility Study}

Show the graphic about domains in a MMORG system\\ Show the
localization of syneight in that model

%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
\subsubsection{Production Domain}

describe and show requirements

%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
\subsubsection{Maintenance Domain}

describe and show requirements

%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
\subsubsection{Gaming Domain}

describe and show requirements

\begin{enumerate}
\item Abstract Requirements
\item Gaming has stable requirements
\item Production/Maintainer domains have unstable requirements
\end{enumerate}

%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
\section{Approach of \SYNEIGHT}
Leading idea: ``Where is \SYNEIGHT: We have an idea of the domain, we
have a very abstract idea of \SYNEIGHT, how do these concepts match?
What is \SYNEIGHT doing in the context of the domain specific services
of a \MMORG.''

Idea. Starting from the domain specific services, we go refine the
concept of \SYNEIGHT subsequently. We give an informal and quite
abstract description of the interface. Then we introduce stepwise
other central forces such as target platforms. And we describe how we
reflected these forces within \SYNEIGHT. The end of the Section should
give the reader a summarizing, lengthy definition of \SYNEIGHT (not
necessarily marked as a formal definition).


\SYNEIGHT covers that and that and leaves that and that issue to the
higher levels.

And why we do this selection. 
\begin{itemize}
\item remember the conclusion of the last section.
\item then we can support other \NVES -- you could say, \SYNEIGHT is
the synchronization essence of an \NVE.
\end{itemize}

%\subsection {\SYNEIGHT Load Distribution}

use this list for argumentation: Load Distribution (LD)
\begin{itemize}
        \item NVEs of a certain size require Clusters
        \item LD more fundamental
        \item Traffic slenderizing
        \item Data Reckoning and Data Bulking
              $\Rightarrow$ Internet (Traffic costs
              really a lot!)\\
              $\Rightarrow$ Good strategies are App/Game
              specific (Domain specific)
	\item Ideally it should be possible to have "Virtual Machine"
\end{itemize}

%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
\section{Alternatives}

%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
\subsection{From Scratch}

One-Time only shot >\_< (KAPUTT)

%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
\subsection{\MMORG Engines}

\begin{itemize}
	\item Domain specific (\_NOT\_FOR\_US\_)
	\item Inferior
	\item List of available engines: Zyta, Butterfly (Ask A.V.M.)
\end{itemize}

Too restrictive. Does not deal with the chellanges. Are not Seamless!

%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
\subsection{Existing Technologies}

\subsubsection{Relational and Object Oriented Data Base Systems}

Not built for Seamless Real Time Cluster systems with OO\\
Client server structure, we are P2P

\subsubsection{Distributed Shared Memory}

Locking?\\
Not built for Seamless Real Time Cluster systems with OO\\
Uncontrolled Growth

\subsubsection{Message Passing Systems and Frameworks}

list multiple solutions: PVM MP Mosix ...\\
easy

Describe 3-4 most known representatives
\begin{itemize}
	\item Not OO
	\item Not Seamless
	\item Intra-process syncronisation
	\item Not uniform
\end{itemize}

% vim:ts=4:sw=4
