%------------------------------------------------------------------------------
% Syneight - A soft-realtime transaction monitor.
% Copyright (C) 2003-2004 The Syneight Group.
%
% TODO: License.
%------------------------------------------------------------------------------

\chapter{Motivation Draft}

%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
\section{What is \SYNEIGHT}
Leading idea: ``Abstract description of \SYNEIGHT''.

Pure prose.

In short:
\begin{description}
\item[Domain independent] define -- use C++ Network Programming,
Schmidt, page 8
\item[Transaction oriented] use the term extended atomic transactions
\item[Soft real-time] define -- Kopetz
\item[Cluster / SMP] SMP: uniform memory, cluster. Clustering for
scalability and availability. We use SMP machines as a nodes since to
keep the number of nodes small -- and thus, the communication
overhead.
\end{description}

Think about ordering of these issues. 

Problem: Not \MMORGS, not \NVES. These two are motivations which we
will discuss in great length in this chapter -- but they are not
defining \SYNEIGHT. Thus, they do not appear in this section.

10 Sentences


%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
\section{Motivating Application Domain}
Leading idea: ``Which domain, why is the domain interesting, which
domain structure.''

Better Structure (convert this section here into this form):
\begin{itemize}
	\item MMORGS -- A new Entertainment World
	\begin{itemize}
		\item Define MMORGS
		\item ``Why MMORGS''
		\item short: the rest of the section
	\end{itemize}
	\item ''what are nves'' with applications
	\item MMORGS in detail
	\begin{itemize}
		\item maybe more detailed definition
		\item MMORGS are a special case of NVEs
		\item industrial interest
  		\item characteristics \& challenges
		\item feasibility study
	\end{itemize}
\end{itemize}

%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
\subsection{\MMORGS -- A new Entertainment World}
Needs a better catchy expression:
\begin{itemize}
	\item A new Entertainment World
	\item The new Chellange
	\item \dots
\end{itemize}

\subsubsection{What are \MMORGS}
Definition \ldots

\subsubsection{The interest in \MMORGS}
\begin{itemize}
	\item Academic Interest
	\begin{itemize}
		\item Subsumes Many general NVE Problems
		\item It is a game:
		\item Incredible wishlist
		\item Tolerant ( The world is not going under when something crashes )
		\item Bottomline: Interesting research ground
	\end{itemize}
	\item Industrial Interest
	\begin{itemize}
		\item Feasibility study
	\end{itemize}
\end{itemize}

%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
\subsection{What are \NVES}

\MMORG is a specialization of \NVE.

based on virtual environments papers

adding networking (cf. overview paper) -- usage of networking in terms
of a, load distribution, and b, geographically dispersed access

read the stuff! /usr/local/data/syneight/literature/nve/survey, read
the papers, check out what's going on.

Also integrate the applications:
\begin{itemize}
        \item Military Simulations
        \item CSCW ( Computer Supported Cooperative Work for example
              Virtual meetings)
        \item Educational
        \item Entertainment
        \item \dots
\end{itemize}

Application Elements? Leave blank in the first run.
\begin{itemize}
        \item X-Reality
        \item Augmented Reality
        \item Simulation
        \item \dots
\end{itemize}

%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
\section{\MMORGS in detail}
Characteristics \& Challenges
\begin{itemize}
	\item maybe more detailed definition
	\item MMORGS are a special case of NVEs
	\item industrial interest
	\item characteristics \& challenges
	\item feasibility study
\end{itemize}

%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
\section{The services of a \MMORG}
Leading Idea: ``OK. So that's the way you can structure such a
biest.''

\textbf{Feasibility Study}

Show the graphic about domains in a MMORG system\\ Show the
localization of syneight in that model

%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
\subsubsection{Production Domain}

describe and show requirements

%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
\subsubsection{Maintenance Domain}

describe and show requirements

%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
\subsubsection{Gaming Domain}

describe and show requirements

\begin{enumerate}
\item Abstract Requirements
\item Gaming has stable requirements
\item Production/Maintainer domains have unstable requirements
\end{enumerate}

%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
\section{Approach of \SYNEIGHT}
Leading idea: ``Where is \SYNEIGHT: We have an idea of the domain, we
have a very abstract idea of \SYNEIGHT, how do these concepts match?
What is \SYNEIGHT doing in the context of the domain specific services
of a \MMORG.''

Idea. Starting from the domain specific services, we go refine the
concept of \SYNEIGHT subsequently. We give an informal and quite
abstract description of the interface. Then we introduce stepwise
other central forces such as target platforms. And we describe how we
reflected these forces within \SYNEIGHT. The end of the Section should
give the reader a summarizing, lengthy definition of \SYNEIGHT (not
necessarily marked as a formal definition).


\SYNEIGHT covers that and that and leaves that and that issue to the
higher levels.

And why we do this selection. 
\begin{itemize}
\item remember the conclusion of the last section.
\item then we can support other \NVES -- you could say, \SYNEIGHT is
the synchronization essence of an \NVE.
\end{itemize}

%\subsection {\SYNEIGHT Load Distribution}

use this list for argumentation: Load Distribution (LD)
\begin{itemize}
        \item NVEs of a certain size require Clusters
        \item LD more fundamental
        \item Traffic slenderizing
        \item Data Reckoning and Data Bulking
              $\Rightarrow$ Internet (Traffic costs
              really a lot!)\\
              $\Rightarrow$ Good strategies are App/Game
              specific (Domain specific)
	\item Ideally it should be possible to have "Virtual Machine"
\end{itemize}

%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
\section{Alternatives}

%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
\subsection{From Scratch}

One-Time only shot >\_< (KAPUTT)

%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
\subsection{\MMORG Engines}

\begin{itemize}
        \item Domain specific (\_NOT\_FOR\_US\_)
        \item Inferior
        \item List of available engines: Zyta, Butterfly (Ask A.V.M.)
\end{itemize}

Too restrictive. Does not deal with the chellanges. Are not Seamless!

%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
\subsection{Existing Technologies}

\subsubsection{Relational and Object Oriented Data Base Systems}

Not built for Seamless Real Time Cluster systems with OO\\
Client server structure, we are P2P

\subsubsection{Distributed Shared Memory}

Locking?\\
Not built for Seamless Real Time Cluster systems with OO\\
Uncontrolled Growth

\subsubsection{Message Passing Systems and Frameworks}

list multiple solutions: PVM MP Mosix ...\\
easy

Describe 3-4 most known representatives
\begin{itemize}
        \item Not OO
        \item Not Seamless
        \item Intra-process syncronisation
        \item Not uniform
\end{itemize}

% vim:ts=4:sw=4
