%------------------------------------------------------------------------------
% Syneight - A soft-realtime transaction monitor.
% Copyright (C) 2003-2004 The Syneight Group.
%
% TODO: License.
%------------------------------------------------------------------------------

\chapter{C++}
\label{cha:C++}


\section{Basics}
\label{sec:basics}


\SYNEIGHT's language of choice will be C++ for reasons that are not
entirely in the scope of this document. For some of them, read
section \vref{sec:goal} in this document. The coding guidelines will base on
the book \emph{Large Scale C++ Software Design} by John Lakos.


\section{Indentation Style}
\label{sec:indentation-style}

All indentation should use tabs in favor for spaces. The reason is
that a mixture of tabs and spaces in a single file can easily produce
faulty indentation and lead to further errors in human reading. The
decision for tabs before spaces is that tabs can display the width
desired.


\section{Compatibility}
\label{sec:compatibility}

I had a dream -- C++ is a uniform language across a large number of
platforms. It is not. It is painful to maintain C++-code in a portable
manner. This section is defining which features of 
\begin{itemize}
\item the C++-language,
\item required tools for the build process,
\item and the used libraries 
\end{itemize}
we require for \SYNEIGHT. After reading this statement, you will think
that it is \textbf{absolutely impossible to state all these
  requirements explicitly.} 
%
Therefore, we adopt the following approach. First, we will list some
central and required features explicitly. Such a list will never
ensure, that \SYNEIGHT is really compiling on a certain platform. On the
other hand, it will rule out platforms which do not satisfy these
requirements.
%
More concretely, we will categorize platforms as follows.
\begin{description}
\item[supported.] \SYNEIGHT is actively supported on these platforms. 
\item[tested.] \SYNEIGHT has been compiled on the given platform and
  its testsuite has been executed successfully. 
\item[broken.] \SYNEIGHT is known not to work on the given platform.
\end{description}

The list of central and required features will be based on other
portable software artifacts. As starting point we will use the
\texttt{boost} feature classification.


\section{Portability Guidelines}
\label{sec:portability-guidelines}

TODO.


%%% Local Variables: 
%%% mode: latex
%%% TeX-master: "standards"
%%% End: 
% vim:ts=4:sw=4
